---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# psychHypergraph

<!-- badges: start -->
<!-- badges: end -->

**psychHypergraph** implements information-theoretic hypergraph analysis for psychometric data, based on Marinazzo et al. (2024) and Possenti et al. (2026). The package provides tools for constructing psychometric hypergraphs, identifying synergistic and redundant higher-order interactions, centrality analysis, clustering, and visualization.

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rheniumyuan/psychHypergraph")
```

## Key Features

- **Hypergraph construction** via `build_hypergraph()`: a complete pipeline that preprocesses data, generates candidate hyperedges, computes O-information, and performs three-step hyperedge validation and selection (permutation test, bootstrap stability assessment, and hierarchical validation) (Possenti et al., 2026).
- **O-information**: an information-theoretic measure proposed that quantifies the balance between redundant and synergistic higher-order statistical dependencies among variables (Rosas et al., 2019).
- **Multiple candidate generation strategies**: implements exhaustive search and clique-based candidate generation strategies.
- **Node degree centrality** via `node_degree()`: calculate node degree centrality for each node.
- **Spectral clustering** via `spectral_clustering()`: perform spectral clustering (similar to `HySC` in python library `hypergraphx (HGX)`) on hypergraphs.
- **Visualization**: `plot()` for hypergraph visualization (based on `HyperG` package), `plot_hyperedge()` for forest plots of O-information estimates, and `plot_node_degree()` for centrality profiles.

## Example

### Constructing a psychometric hypergraph

```{r example}
library(psychHypergraph)

# Load example data
data(bfi, package = "psych")

# Build a psychometric hypergraph using clique-based candidate generation
hg <- build_hypergraph(
  data = bfi[, 1:25],
  cor_method = "pearson",
  k = 3:4,
  candidate_method = "clique",
  n_perm = 100, # Reduced for faster example; use 1000+ for real analysis
  n_boot = 100,
  seed = 123,
  parallel = FALSE
)

# Inspect results
print(hg)
```

### Analyzing the structure of the hypergraph

```{r}
# Node degree centrality
plot_node_degree(hg, graph_type = c("synergistic","redundant"), style = "both")

# Spectral clustering
clustering <- spectral_clustering(hg, K = 5)
print(clustering)
```

### Visualizing the hypergraph

```{r plot}
# Visualize the hypergraph
plot(hg, top_n = 30, type = "full", vertex.size = 20, groups = clustering$membership)

# visualize hyperedges
plot_hyperedge(hg, show_edge_label = FALSE, significant_only = TRUE)
```

## References

- Marinazzo, D., Van Roozendaal, J., Rosas, F. E., Stella, M., Comolatti, R., Colenbier, N., Stramaglia, S., & Rosseel, Y. (2024). An information-theoretic approach to build hypergraphs in psychometrics. *Behavior Research Methods*, *56*(7), 8057–8079. https://doi.org/10.3758/s13428-024-02471-8
- Possenti, F., Girelli, L., Tieri, P., & Petti, M. (2026). Multiplex Hypergraph Modeling of Higher Order Structures in Psychometric Networks (Version 1). arXiv. https://doi.org/10.48550/ARXIV.2604.22744
- Rosas, F. E., Mediano, P. A. M., Gastpar, M., & Jensen, H. J. (2019). Quantifying high-order interdependencies via multivariate extensions of the mutual information. *Physical Review E*, *100*(3), 032305. https://doi.org/10.1103/PhysRevE.100.032305
